---
title: "Colocalisation"
author: "MW"
date: "`r Sys.Date()`"
output: html_document
---
## Load libraries
Using Bayesian test for colocalisation from https://journals.plos.org/plosgenetics/article?id=10.1371/journal.pgen.1004383

```{r}
library(TwoSampleMR)
library(tidyverse)
library(coloc)
```

## Read in data
Coloc also needs proportion case:control (s) for the outcome, in this case prostate cancer, as input.
```{r}
trait <- "SPINT2"
exp_dat <- read_exposure_data(paste0(file="exposure_coloc_input_",trait,".txt"), sep="\t")
out_dat <- read_outcome_data(paste0(file="outcome_coloc_input_",trait,".txt"), sep="\t")
s <- 	0.564437729 
```

## Harmonise data
Use harmonise_data function in TwoSampleMR 
```{r}
dat <- harmonise_data(exp_dat,out_dat, action=1) #used TwoSampleMR to harmonise data

```

## Create outcome dataset
```{r}
SNPs <- dat$SNP
outcomebeta <- dat$beta.outcome
names(outcomebeta)<-SNPs
outcomevarbeta <- dat$se.outcome^2
names(outcomevarbeta)<-SNPs
pos <- dat$pos.outcome
names(pos)<- SNPs
type <- "cc" #casecontrol


outcomedataset <- list(snp=SNPs, beta=outcomebeta, varbeta=outcomevarbeta, position=pos, type=type, s=s)
```

## Check and plot outcome dataset
```{r}
check_dataset(outcomedataset)
plot_dataset(outcomedataset)
```

## Create exposure dataset
```{r}
SNPs <- dat$SNP
traitbeta <- dat$beta.exposure
names(traitbeta)<-SNPs
traitvarbeta <- dat$se.exposure^2
names(traitvarbeta)<-SNPs
pos <- dat$pos.outcome
names(pos)<- SNPs
type <- "quant"
sdY <- 1

traitdataset <- list(snp=SNPs, beta=traitbeta, varbeta=traitvarbeta, position=pos, type=type, sdY=sdY)

```

## Check and plot exposure dataset
```{r}
check_dataset(traitdataset)
plot_dataset(traitdataset)
```

## Run coloc
```{r}
my.res <- coloc.abf(traitdataset,outcomedataset, MAF = NULL, p1 = 1e-04, p2 = 1e-04, p12 = 1e-05)
```

## Write data
```{r}
results <- as.data.frame(as.list(my.res$summary))
results$trait <- trait
results$cancer <- "prostate"
write_tsv(results, paste0(file=trait,"_coloc_results.txt"))
```

